---
title: "Data Prep & Descriptive Statistics"
output: html_notebook
---

```{r}
library(tidyverse)
library(RCurl)
```
#Importing the data
```{r}
train <- read.csv("https://raw.githubusercontent.com/crisajose/CIND-820-Big-Data-Analytics-Project/main/train.csv", header=T)

```

# number of observations and number of variables in dataset
```{r}
dim(train)
```

# displaying structure and a subset of values of dataset
```{r}
glimpse(train)
```

#summary of predictive variable(Response), BMI, Weight & Height.
```{r}
summary(train$Response)
summary(train$BMI)
summary(train$Wt)
summary(train$Ht)
```

#Checking missing values present in dataset
```{r}
colSums(is.na(train))
```
#Find columns with missing values
```{r}
Missingcol<-colnames(train)[colSums(is.na(train)) > 0]
Missingcol

```



#11 out of 13 columns having NA have very high missing values

# Checking Correlation between BMI, Wt and Ht
```{r}
cor(train$BMI,train$Wt)
```
```{r}
cor(train$BMI,train$Ht) 
```

```{r}
cor(train$Ht,train$Wt)
```

We find that BMI is highly correlated with Wt.

```{r}
library(corrplot)
newclean<- train[,c(10,11,12)]
clean_corr <-  as.matrix(cor(newclean))
corrplot(clean_corr, method="number")
corrplot(clean_corr, method="pie")
```


#Examine variable Response

```{r}
library(ggplot2)
responseg <- ggplot(data = train, aes(Response)) + geom_bar()
responseg
#Response values for 3 and 4 are very rare. Response 8 is by far the most common value
```

#BMI seem to have an impact on the response variable based on the graph



#Response value 8 has lower value of BMI compared to response value 1.Therefore we see that BMI has a strong relationship with Response variable
```{r}
# Boxplot Response ~ BMI
ggplot(train, aes(x=factor(Response), y=BMI)) + ggtitle("Boxplot Response ~ BMI") + 
  geom_boxplot(colour="blue")+  theme(axis.text.x=element_text(angle=90,hjust=1))
```

#we see that Ht does not have a strong relationship with Response variable and does not affect it as much
```{r}
# Boxplot Response ~ Ht
ggplot(train, aes(x=factor(Response), y=Ht)) + ggtitle("Boxplot Response ~ Ht") + 
  geom_boxplot(colour="red")+  theme(axis.text.x=element_text(angle=90,hjust=1))
```


#we see that Age has a relationship with Response variable
```{r}
ggplot(train, aes(x=factor(Response), y=Ins_Age)) + ggtitle("Boxplot Response ~ Ins_Age") + 
  geom_boxplot(colour="blue")+  theme(axis.text.x=element_text(angle=90,hjust=1))
```

#We can see that distribution of response variable is dependent on Family_Hist_1
```{r}
# histogram of Response on Family_Hist_1
  
  ggplot(train, aes(x=Response)) + ggtitle("Histogram  Response with Family_Hist_1 values") +
    geom_histogram(position="identity", colour="black", alpha=0.2, bins = 10)+
    facet_grid(. ~ Family_Hist_1)
```

