---
title: "Data Prep & Descriptive Statistics"
output: html_notebook
---

```{r}
library(tidyverse)
library(RCurl)
```
#Importing the data
```{r}
train <- read.csv("https://raw.githubusercontent.com/crisajose/CIND-820-Big-Data-Analytics-Project/main/train.csv", header=T)

```

# number of observations and number of variables in dataset
```{r}
dim(train)
```

# displaying structure and a subset of values of dataset
```{r}
glimpse(train)
```

#summary of predictive variable(Response), BMI, Weight & Height.
```{r}
summary(train$Response)
summary(train$BMI)
summary(train$Wt)
summary(train$Ht)
```

#Checking missing values present in dataset
```{r}
colSums(is.na(train))
```
#Find columns with missing values
```{r}
Missingcol<-colnames(train)[colSums(is.na(train)) > 0]
Missingcol

```



#11 out of 13 columns having NA have very high missing values

# Checking Correlation between BMI, Wt and Ht
```{r}
cor(train$BMI,train$Wt)
```
```{r}
cor(train$BMI,train$Ht) 
```

```{r}
cor(train$Ht,train$Wt)
```

We find that BMI is highly correlated with Wt.

```{r}
library(corrplot)
newclean<- train[,c(10,11,12)]
clean_corr <-  as.matrix(cor(newclean))
corrplot(clean_corr, method="number")
corrplot(clean_corr, method="pie")
```


#Examine variable Response

```{r}
library(ggplot2)
responseg <- ggplot(data = train, aes(Response)) + geom_bar()
responseg
#Response values for 3 and 4 are very rare. Response 8 is by far the most common value
```

#BMI seem to have an impact on the response variable based on the graph



#Response value 8 has lower value of BMI compared to response value 1.Therefore we see that BMI has a strong relationship with Response variable
```{r}
# Boxplot Response ~ BMI
ggplot(train, aes(x=factor(Response), y=BMI)) + ggtitle("Boxplot Response ~ BMI") + 
  geom_boxplot(colour="blue")+  theme(axis.text.x=element_text(angle=90,hjust=1))
```

#we see that Ht does not have a strong relationship with Response variable and does not affect it as much
```{r}
# Boxplot Response ~ Ht
ggplot(train, aes(x=factor(Response), y=Ht)) + ggtitle("Boxplot Response ~ Ht") + 
  geom_boxplot(colour="red")+  theme(axis.text.x=element_text(angle=90,hjust=1))
```


#we see that Age has a relationship with Response variable
```{r}
ggplot(train, aes(x=factor(Response), y=Ins_Age)) + ggtitle("Boxplot Response ~ Ins_Age") + 
  geom_boxplot(colour="blue")+  theme(axis.text.x=element_text(angle=90,hjust=1))
```

#We can see that distribution of response variable is dependent on Family_Hist_1
```{r}
# histogram of Response on Family_Hist_1
  
  ggplot(train, aes(x=Response)) + ggtitle("Histogram  Response with Family_Hist_1 values") +
    geom_histogram(position="identity", colour="black", alpha=0.2, bins = 10)+
    facet_grid(. ~ Family_Hist_1)
```

Convert Response–our target variable–to a factor

```{r}
train$Response <- as.factor(train$Response)
```


Check for NA
```{r}
sapply(train, function(x) sum(is.na(x)) )
```

Delete all rows with NA
```{r}
cleantrain <- na.omit(train)

```

DECISION TREE
```{r}
library(rpart)
```
```{r}
dtmodel <- rpart(train$Response ~., data=train, method = "class")


dtmodel
```
```{r}
install.packages("rpart.plot")
library(rpart.plot)
rpart.plot(dtmodel, type=1, extra = 102)
```

#Prepare test set

```{r}
#test <- read.csv("https://michaschweizer.github.io/datasets/test.csv", na.strings = "")
test <- read.csv("https://raw.githubusercontent.com/crisajose/CIND-820-Big-Data-Analytics-Project/main/test.csv", header=T)

#test <- subset(test, select = -c(Employment_Info_6, Insurance_History_5,Family_Hist_2, Family_Hist_3, Family_Hist_4, Family_Hist_5, Medical_History_1,Medical_History_10,Medical_History_15,Medical_History_24, Medical_History_25,Medical_History_32) )

sapply(test, function(x) sum(is.na(x)) )
```


```{r}
test[!complete.cases(test),'Employment_Info_1'] <- median(test$Employment_Info_1, na.rm = T)

```

#decision tree to predict the Response values for the test set

```{r}
dttest <- predict(dtmodel, test, type = "class")
```


```{r}
predictions <- data.frame(test$Id)
predictions$Response <- dttest
names(predictions) <- c('Id', 'Response')

write.csv(predictions, 'myprediction_tree.csv', row.names = F)
```

